Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours

### 2015-02-19 Publications/2015-NeuroImage/03-introduction.tex
finishing changes

In this work we propose a hybrid approach to nonlinear registration of anatomically correct
  and precise surfaces to warped and low-resolution \gls*{dmri} of the brain.
Currently, we are witnessing an uprising demand for such algorithms in the challenging context
  of extracting and characterizing the structural connectivity networks of the
  human brain \citep{craddock_imaging_2013}.
Our registration method is demonstrated here as a solution to correct for the susceptibility
  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
\todo[inline]{talk about the distortion and cite SDCs}

A fundamental antecedent of the present approach is found in \citep{greve_accurate_2009}.
They proposed a registration method to align \gls*{t1} and \gls*{dmri} images
  of the brain using a linear transformation model.
Similarly, we exploit the precise structures that can be extracted with
  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}),
  and then map these surfaces into warped data, in which the scale of the
  imaged structures is in the range of the image resolution or above.
Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
First, we use a nonlinear deformation model that enables the correction of the
  susceptibility-derived distortions \citep{jezzard_correction_1995}.
Second, our matching criteria does not necessarily need for clear intensity
  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
  active contours without edges \citep{chan_active_2001}.
The hypothesis underpinning our research is that, in those situations,
  image registration can be reliably performed by searching for homogeneous
  regions in one image corresponding to a prior knowledge of the shape of objects in
  the moving coordinate system.

>>>>

In this work we propose a hybrid approach to nonlinear registration of anatomically correct
  and precise surfaces to warped and low-resolution \gls*{dmri} of the brain.
Currently, we are witnessing an uprising demand for such algorithms in the quest
  of extracting and characterizing the networks of structural connectivity in the
  human brain \citep{craddock_imaging_2013}.
Our registration method is demonstrated here as a solution to correct for the susceptibility
  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
The precise segmentation of \gls*{dmri} data is also crucial in such applications, since the
  definition of the nodes in the network is generally derived from a parcellation performed in
  an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.

\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
In order to quickly probe  in many different orientations the diffusion process within the living
  brain, these sequences minimize the bandwidth of the \gls*{pe} direction.
This low bandwith originates the artifactual displacement of locations presenting small deviations
  from the main magnetic field $B_0$.
The principal source of these deviations, $\Delta B_0$, is the discontinuity of magnetic
  susceptibility at tissue interfaces.
Tissue/air boundaries show large steps of susceptibility and this artifact is
  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
  bones in both hemispheres.
The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
  \citep{jezzard_correction_1995}:

  \begin{equation}
  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE}.
  \label{eq:fieldmap}
  \end{equation}

where $\vec{r}$ is the location in the reference domain as in \eqref{eq:transform},
  $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
  vector along \gls*{pe}.
Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
The first solution was acquiring extra images to estimate $\Delta B_0$ using field mapping
  techniques \citep{jezzard_correction_1995,andersson_modeling_2001}.
A second set of methods aquired extra images reproducing the same \gls*{epi} scheme but switching
  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient
  increments in the same axis \citep{cordes_geometric_2000,holland_efficient_2010}.
\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \gls*{dmri}
  data versus a \gls*{t2} image.
Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
  three-dimensional problem.
All methods require extra \gls*{mr} aquisitions. Thus, a tipical protocol for the extraction
  of structural connectivity comprehends \gls*{dmri}, \gls*{t1}, and the acquisition demanded
  by the correction method of choice.

Our method is related to the \gls*{t2b} methods as we also render the correction as a
  registration problem.
Conversely, the method does not require any of the previously mentioned extra acquisitions.
A fundamental antecedent to the present approach is found in \citep{greve_accurate_2009}.
They proposed a registration method to align \gls*{t1} and \gls*{dmri} images
  of the brain using a linear transformation model.
Similarly, we exploit the precise structures that can be extracted with
  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}),
  and then map these surfaces into warped data, in which the scale of the
  imaged structures is in the range of the image resolution or above.
Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
First, we use a nonlinear deformation model that enables the correction of the
  susceptibility-derived distortions \citep{jezzard_correction_1995}.
Second, our matching criteria does not necessarily need for clear intensity
  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
  active contours without edges \citep{chan_active_2001}.
The hypothesis underpinning our research is that, in those situations,
  image registration can be reliably performed by searching for homogeneous
  regions in one image corresponding to a prior knowledge of the shape of objects in
  the moving coordinate system.

### 2015-02-19 Publications/2015-NeuroImage/04-methods.tex
finishing changes

An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000}.
Shortly, \cite{yezzi_variational_2001} presented the first method including a full solution to
  the registration, where the energy functional is defined simultaneously in the target
  and reference images with an affine transformation supporting the coordinates mapping.
\cite{vemuri_joint_2003} proposed an atlas-based registration using a level sets
  framework and only one \gls*{pde} for first.
The principal application of these techniques was the image registration between timesteps in
  a time-series or between different images \citep{paragios_level_2003}.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2001}
  to nonlinear registration with a free deformation field.
\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration-segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
Finally, \cite{gorthi_active_2011} presented a comprehensive generalization of the
  existing methodologies based on the level sets approximation in the application of
  atlas-based segmentation.
A recent solution of the segmentation-registration problem close to the one proposed here was
  presented by \cite{guyader_combined_2011}.
Our method differs on two key features:
  it allows multiple active contours for segmentation, and the regularization strategy.
Whereas \cite{guyader_combined_2011} proposed a nonlinear elasticity smoother, we rely on
  the smoothing effects of the deformation field and the anisotropic regularization
  proposed by \cite{nagel_investigation_1986}.

@@

  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between tissues $l$ and $m$.
Then, the \gls*{map} criterium is applied to find the mapping estimate $\hat{U}$
  that aligns $\Gamma$ into $R$:

  \begin{align}
  \hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid R,\Gamma_{l,m}) \right\} \notag\\
   &= \underset{U}{\argmax} \left\{ P(R \mid U,\Gamma_{l,m})\, P(U) \right\}.
  \label{eq:map_u}
  \end{align}

@@

  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} of the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions:

  \begin{equation}
  P\left( \vec{f}' \mid U,\Omega_l \right) = \mathcal{N} \left( \vec{f}' \mid \Theta_l \right),
  \label{eq:likelihood}
  \end{equation}

 	where $\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$ are the
 	descriptors of region $\Omega_l$.
We then insert \eqref{eq:likelihood} into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

@@

We choose a Thikonov regularization prior as follows:

  \begin{equation*}
  P(U) = \underset{\vec{r}}{\prod} p(u(\vec{r})) =
  \underset{\vec{r}}{\prod} p_0(u(\vec{r})) \, p_1(u(\vec{r})),
  \end{equation*}

  with further:

  \begin{align*}
  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, B^{-1}),
  \end{align*}

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
  surface(s).
In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}.
The transformation from the structural space into the coordinate space of the
  target image is achieved through a dense deformation field $u(\vec{r})$, such that:

  \begin{equation}
  \vec{v}_i' = U\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i),
  \label{eq:nodes_tfm}
  \end{equation}

  where $U$ is defined in \eqref{eq:transform}.
Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:

@@

In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3 ($B_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
  $\psi$ by the actual kernel function, the transformation writes:

  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      B_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}

@@

{\color{red}
\paragraph*{Convergence evidence and settings}
 discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
\paragraph*{Efficient field densification}
As long as the dense deformation field is iteratively interpolated
from the same set of control points that define the $L-1$ contours,
it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
sparse matrix for fast densification. As well, we achieve a
diffeomorphic transform by}

@@

\begin{figure}
\caption{1) The prior surfaces are extracted from an undistorted and high-resolution dataset
	  (e.g. \autoref{fig:phantom}). 2) A plausible, synthetic distortion $U^{-1}_{true}$ is generated.
	3) The low-resolution (target) images are warped using the synthetic distortion.
	4) Our method is run, obtaining a $\hat{U}_{tst}$.
	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
\end{figure}

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{figures/figure02}
	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
	  outer crust resembling the cortical folding (left).
	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
	B. Visual assessment of the results on the low resolution sets:
	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
	  and ``box'' at (bottom-right).
	In yellow color, the recovered contours after registration are represented.
	Our method showed high accuracy, as they are overlapping the ground truth surfaces
	  depicted in green.
	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
	  clustering methods, but it is successfully segmented with our method.
	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
	  consistently below the image resolution.
	  }\label{fig:phantom}
\end{figure*}

@@

In this case, we derived the deformation $U_{true}^{-1}$ from the field maps released with
  the corresponding packages of each dataset from the \gls*{hcp},
  mimicking the real distortions, using a derivation of our previous work
  \citep{esteban_simulationbased_2014}.
After warping the original \gls*{dmri} with $U_{true}^{-1}$, we computed the \gls*{dti} and
  the derived scalars (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.

>>>>

An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000}.
Shortly, \cite{yezzi_variational_2001} presented the first method including a full solution to
  the registration, where the energy functional is defined simultaneously in the target
  and reference images with an affine transformation supporting the coordinates mapping.
\cite{vemuri_joint_2003} proposed an atlas-based registration using a level sets
  framework and only one \gls*{pde} for first.
The principal application of these techniques was the image registration between timesteps in
  a time-series or between different images \citep{paragios_level_2003}.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2001}
  to nonlinear registration with a free deformation field.
\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration-segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
Finally, \cite{gorthi_active_2011} presented a comprehensive generalization of the
  existing methodologies based on the level sets approximation in the application of
  atlas-based segmentation.
Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method close to the one proposed here.
Our method differs on two key features:
  it allows multiple active contours for segmentation, and the regularization strategy.
Whereas \cite{guyader_combined_2011} proposed a nonlinear elasticity smoother, we rely on
  the smoothing effects of the deformation field and the anisotropic regularization
  proposed by \cite{nagel_investigation_1986}.

@@

  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between competing regions $\Omega_l$ and $\Omega_m$.
The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
  \citep{bishop_pattern_2006}, and aligns $\Gamma$ into $R$.

@@

  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

@@

The smoothness of the resulting displacements field is induced by a Thikonov regularization
  prior:

  \begin{align*}
  P(U) = \underset{\vec{r}}{\prod}\, p(u(\vec{r})) &=
  \underset{\vec{r}}{\prod}\, p_0(u(\vec{r})) \, p_1(u(\vec{r})), \\
  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, B^{-1}),
  \end{align*}

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
  surface(s).
In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}.
The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
  $u(\vec{r})$, such that:

  \begin{equation}
  \vec{v}_i' = \hat{U}\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i).
  \label{eq:nodes_tfm}
  \end{equation}

Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:

@@

In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3 ($\beta_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
  $\psi$ by the actual kernel function, the transformation writes:

  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}

@@

{\color{red} \paragraph*{Evidencing convergence and settings}
discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions,
multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figures/figure02}
\caption{Experimental workflow applied on real data from the \gls*{hcp}.
  1) The prior surfaces are extracted from the anatomical reference (\gls*{t1} image).
	2) To operate as ground truth, we generate plausible-synthetic distortion $U^{-1}_{true}$
	  from the fieldmap using \eqref{eq:fieldmap}.
	3) The \gls*{dmri} data are warped using $U^{-1}_{true}$ to reproduce the effects of real
	  susceptibility-derived distortions.
	Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
		stacked to feed the multivariate input required by our algorithm.
	4) Registration is run, obtaining a $U_{test} = \hat{U}_{true}$, an estimation of
	  the ground-truth deformation $U_{true}$.
	  A cross-comparison methodology is also applied, to obtain a competing $U_{cc}$.
	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{figures/figure03}
	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
	  outer crust resembling the cortical folding (left).
	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
	B. Visual assessment of the results on the low resolution sets:
	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
	  and ``box'' at (bottom-right).
	In yellow color, the recovered contours after registration are represented.
	Our method showed high accuracy, as they are overlapping the ground truth surfaces
	  depicted in green.
	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
	  clustering methods, but it is successfully segmented with our method.
	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
	  consistently below the image resolution.
	  }\label{fig:phantom}
\end{figure*}

@@

In this case, we derived the deformation $U_{true}^{-1}$ from the field maps released with
  the corresponding packages of each dataset from the \gls*{hcp} applying \eqref{eq:fieldmap},
  mimicking the real distortions, using a derivation of our previous work
  \citep{esteban_simulationbased_2014}.
After warping the original \gls*{dmri} with $U_{true}^{-1}$, we computed the \gls*{dti} and
  the derived scalars (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.

### 2015-02-19 Publications/2015-NeuroImage/05-results.tex
finishing changes

\subsection*{Proof of concept on digital phantoms}
\label{sec:results_phantom}
Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
  we collected quantitative results on the digital phantoms.
To warp the simulated datasets, we generated 150 realizations of random and smooth displacement
  fields $U^{-1}_{true}$.
Smoothness is achieved using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

@@

>>>>

\subsection*{Proof of concept on digital phantoms}
\label{sec:results_phantom}
Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
  we collected quantitative results on the digital phantoms.
To warp the simulated datasets, we randomly generated 150 different realizations of
  the distortion $U^{-1}_{true}$.
Smoothness is introduced using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{figures/figure05.pdf}
	\caption{Results on real data}\label{fig:results_real}
\end{figure*}

### 2015-02-19 Publications/2015-NeuroImage/06-discussion.tex
finishing changes

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
We propose a joint registration and segmentation algorithm designed for biomedical
  image analyses that are affected by all or some of the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).
The registration method performs the segmentation of the target image in homogeneous
  and competing regions defined by explicit active contours.
We chose active contours without edges as we were interested on segmenting homogeneous
  regions of diffusion properties within the brain.
Finally, a B-Spline deformation field supports and keeps track of the evolution of the
  active contours.
Iteratively, the active contours are updated following the instant shape gradients
  \citep{herbulot_segmentation_2006} with a gradient descent strategy.
In section \nameref{sec:methods_background}, we framed our method within the current
  state-of-art and defined the novelty of our approach.

To support our claim that images can be simultaneously segmented and registered to a
  reference coordinate system, we conducted evaluation experiments on digital
  simulated phantoms and real datasets.

The algorithm is designed for the proposed application on correcting distortions of
  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
  networks.
As a consequence, our approach relies on the availability of precise segmentations or
  surfaces of the objects that are to be segmented on the target volume(s).
Connectome extraction protocols generally include the acquisition of \gls*{t1} images
  to provide with prior information about anatomy with high accuracy.
\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
the EPI distortions into regularization}
Our method solves in a single step a joint problem usually addressed in two steps.
The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data (i.e.
  \cite{greve_accurate_2009}), whereas the second corrects for the nonlinear distortions
  derived from the inhomogenety of magnetic susceptibility across tissues
  \citep{jezzard_correction_1995}.

>>>>

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
We present a variational framework using to simultaneously segment and
  register 3D \gls*{dmri} of the human brain using the corresponding-anatomical image
  as reference.
The aproach is valid for image processing applications in which a precise segmentation is
  available (from an atlas or corresponding anatomical data) and match any or all
  the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).

The algorithm is designed for the proposed application on correcting distortions of
  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
  networks.
As a consequence, our approach relies on the availability of precise segmentations or
  surfaces of the objects that are to be segmented on the target volume(s).
Connectome extraction protocols generally include the acquisition of \gls*{t1} images
  to provide with prior information about anatomy with high accuracy.
Our method solves in a single step a joint problem usually addressed in two steps.
The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data (i.e.
  \cite{greve_accurate_2009}), whereas the second corrects for the nonlinear distortions
  derived from the inhomogenety of magnetic susceptibility across tissues
  \citep{jezzard_correction_1995}.

In section \nameref{sec:methods_background}, we framed our method within the current
  state-of-art and defined the novelty of our approach.
The registration method performs the segmentation of the target image in several
  competing regions defined explicitly by their limiting surfaces.
We chose active contours without edges as we were interested in the homogeneous
  regions of diffusion properties within the brain.
Finally, a B-Spline deformation field supports and keeps track of the evolution of the
  active contours.
Iteratively, the active contours are updated following the instant shape gradients
  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
To support our claim that images can be simultaneously segmented and registered to a
  reference coordinate system, we conducted evaluation experiments on digital
  simulated phantoms and real datasets.

\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
the EPI distortions into regularization}

### 2015-02-19 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
finishing changes

\subsection*{Instant shape priors}
\label{app:shape_priors}
Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
  $\Omega = \Omega_l \cup \Omega_m$ splitted in two regions $l$ and
  $m$, and $\Gamma_{l,m}$ a closed boundary between them
  as defined in \nameref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:

@@

  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
Assuming that the region descriptors $\Theta_l$ vary slowly enough, we can consider
  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:

@@

  In our formulation, $F(\vec{r})$ is the difference on
    sample likelihood between the inner ($\Omega_l$) and the outer ($\Omega_m$) regions.
  Therefore, we substitute:

@@

>>>>

\subsection*{Instant velocity of contours}
\label{app:shape_priors}
The computation of gradients at the locations of the active contours in the
  instant $t$ is based on the work of \cite{herbulot_segmentation_2006}.
Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
  $\Omega = \Omega_l \cup \Omega_m$ splitted in two regions $l$ and
  $m$, and $\Gamma_{l,m}$ a closed boundary between them
  as defined in \nameref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:

@@

  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
Assuming that the region descriptors $\{\mu_l, \Sigma_l\}$ vary slowly enough, we can consider
  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:

@@

  As inner ($\Omega_l$) and outer ($\Omega_m$) regions are competing and introducing
    the energy functional \eqref{eq:energy}:

@@

\subsection*{Efficient field densification}
As long as the dense deformation field is iteratively interpolated
from the same set of control points that define the $L-1$ contours,
it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
sparse matrix for fast densification. As well, we achieve a
diffeomorphic transform by
